Research Proposal

### 2012-03-14 dissertation-proposal/research-proposal.tex
2nd draft

Multimethods play a significant role in Clojure, and a satisfactory
type system for Clojure should understand them to some degree.
There is some experience statically typing multimethods
\cite{Millstein02modularstatically}
, but this is a challenge and will probably not be completed in the
timeframe proposed.

Clojure's core library includes variable-arity functions which
are not easily typed with current static type systems. 
Providing satisfactory types for functions 
like Clojure's 'map', 'filter', and 'reduce' require
support for non-uniform variable arity polymorphism. 
Strickland, Tobin-Hochstadt and Felleisen \cite{Strickland:2009:PVP:1532974.1532978}.
describe their approach for non-uniform variable arity polymorphism, as used
in Typed Racket. It is not clear if including a similar system is 
achievable in the given timeframe, but would be a very useful feature.

@@

Ensuring vigorous type safety is an important aspect of Typed Scheme 
\cite{TypedScheme:2010} ,
especially when interacting between untyped and typed modules.
I do not expect to concentrate on every combination of cross-module
interaction. In particular, safely using typed code from untyped code
is not important for my needs.

@@

\begin{description}
\item[0.1]
	\begin{itemize}
	\item Concrete types (no type variables)
	\item Union types
	\item Fixed arity function types
	\item Typed deftype
	\item Uniform variable-arity function types
	\item Type checking without refinements or occurance typing
	\end{itemize}
\item[0.2]
	\begin{itemize}
	\item Occurance typing for type inference
	\item Type variables (polymorphic types)
	\end{itemize}
\item[0.3]
	\begin{itemize}
	\item Typed defrecord
	\item All Types annotated for clojure.core library
	\end{itemize}
\item[0.4]
	\begin{itemize}
	\item Manage interactions between typed namespaces
	\end{itemize}
\item[0.5]
	\begin{itemize}
	\end{itemize}
\item[0.6]
	\begin{itemize}
	\item Mutable reference types
	\item Non-uniform variable arity polymorphism
	\end{itemize}

>>>>

Multimethods play a significant role in Clojure, and a satisfactory
type system for Clojure should understand them to some degree.
There is some experience statically typing multimethods
\cite{Millstein02modularstatically}
, but this is a challenge and may not be addressed in the
timeframe proposed.

Clojure's core library includes variable-arity functions which
are not easily typed with current static type systems. 
Providing satisfactory types for functions 
like Clojure's 'map', 'filter', and 'reduce' require
support for non-uniform variable arity polymorphism. 
Strickland, Tobin-Hochstadt and Felleisen \cite{Strickland:2009:PVP:1532974.1532978}.
describe their approach for non-uniform variable arity polymorphism, as used
in Typed Racket. 
This would be a very useful feature, but may be a challenge to implement.

@@

Ensuring vigorous type safety is an important aspect of Typed Scheme 
\cite{TypedScheme:2010} ,
especially when interacting between untyped and typed modules.
I do not expect to concentrate on every combination of cross-module
interaction. In particular, safely using typed code from untyped code
is not a priority.

@@

The milestones for this project are broken into release milestones
for the prototype library.

My goal is to finish all features listed in releases below
and to begin work on the challenges, listed last.

\begin{description}
\item[0.1]
	\begin{itemize}
	\item Concrete types (no type variables)
	\item Union types
	\item Fixed arity function types
	\item Typed deftype (class definitions)
	\item Uniform variable-arity function types
	\item Type checking without refinements or occurance typing
	\item Enough functions annotated from clojure.core for proof-of-concept
	\end{itemize}
\item[0.2]
	\begin{itemize}
	\item Typed defrecord (record structures)
	\item Type variables (polymorphic types)
	\item Most functions annotated from clojure.core library
	\end{itemize}
\item[Challenges]
	\begin{itemize}
	\item Devise strategy for type inference (eg. occurence typing)
	\item Manage interactions between typed namespaces
	\item Mutable reference types
	\item Non-uniform variable arity polymorphism
	\item Typing Multimethods
	\end{itemize}
